---
title: "Hotel Review Analysis"
author: "Viktor Szabo"
output: rmarkdown::github_document
---

In this exercise, I use Natural Language Processing (NLP) to predict whether a hotel rating is positive (greater than 3 stars) based on Tripadvisor user reviews. Using various statistical methods - like CART and random forest - I managed to improve the baseline prediction accuracy (0.69) to 0.83.

## Read in the data

Source: https://www.kaggle.com/andrewmvd/trip-advisor-hotel-reviews

```{r}
reviews = read.csv("tripadvisor_hotel_reviews.csv", stringsAsFactors=FALSE)
reviews$Positive = as.factor(reviews$Rating > 3)
str(reviews)
table(reviews$Positive)
```

Calculate baseline accuracy based on most frequent value:

```{r}
max(table(reviews$Positive)) / nrow(reviews)
```

## Create corpus

```{r}
library(tm)
library(SnowballC)
corpus = VCorpus(VectorSource(reviews$Review))
corpus[[1]]$content
```

Transform text:

```{r}
# Convert to lower-case
corpus = tm_map(corpus, content_transformer(tolower))
# Remove punctuation
corpus = tm_map(corpus, removePunctuation)
# Remove stopwords and hotel
corpus = tm_map(corpus, removeWords, c("hotel", stopwords("english")))
# Stem document 
corpus = tm_map(corpus, stemDocument)
corpus[[1]]$content
```

## Create matrix

```{r}
frequencies = DocumentTermMatrix(corpus)
sparse = removeSparseTerms(frequencies, 0.95)
reviewSparse = as.data.frame(as.matrix(sparse))
# Make all variable names R-friendly
colnames(reviewSparse) = make.names(colnames(reviewSparse))
# Add dependent variable
reviewSparse$PositiveRating = reviews$Positive
```

## Split the data

```{r}
library(caTools)
set.seed(123)
split = sample.split(reviewSparse$PositiveRating, SplitRatio = 0.7)
trainSparse = subset(reviewSparse, split==TRUE)
testSparse = subset(reviewSparse, split==FALSE)
```

## CART model

```{r}
library(rpart)
library(rpart.plot)
reviewCART = rpart(PositiveRating ~ ., data=trainSparse, method="class", minbucket = 15)
prp(reviewCART)
```

Generate confusion matrix

```{r}
predictCART = predict(reviewCART, newdata=testSparse, type="class")
confusionMatrix = table(testSparse$PositiveRating, predictCART)
confusionMatrix
```

Calculate accuracy of CART model:

```{r}
sum(diag(confusionMatrix)) / nrow(testSparse)
```

## Random Forests

```{r message=FALSE}
library(randomForest)
set.seed(1)
reviewRF = randomForest(PositiveRating ~ ., data=trainSparse)
predictRF = predict(reviewRF, newdata=testSparse)
```

Calculate accuracy from confusion matrix:

```{r}
confusionMatrixRF = table(testSparse$PositiveRating, predictRF)
confusionMatrixRF
sum(diag(confusionMatrixRF)) / nrow(testSparse)
```