---
title: "Heart Failure Prediction"
author: "Viktor Szabo"
output: rmarkdown::github_document
---

In this analysis, I make an attempt to predict heart rate failures based on various medical data. Using various statistical methods - like CART and random forest - I managed to improve the baseline accuracy (0.68) to 0.79.

## Read in the data

Source: https://www.kaggle.com/andrewmvd/heart-failure-clinical-data

```{r}
heart = read.csv("heart_failure_clinical_records_dataset.csv")
heart$DEATH_EVENT = as.factor(heart$DEATH_EVENT)
str(heart)
```
## Split the data

```{r}
library(caTools)
set.seed(1)
spl = sample.split(heart$DEATH_EVENT, SplitRatio = 0.7)
Train = subset(heart, spl==TRUE)
Test = subset(heart, spl==FALSE)
```

Calculate baseline accuracy based on most frequent value:

```{r}
max(table(heart$DEATH_EVENT)) / nrow(heart)
```

## CART model

```{r}
library(rpart)
library(rpart.plot)
HeartTree = rpart(DEATH_EVENT ~ ., data = Train, method="class", minbucket=5)
prp(HeartTree)
```

Generate confusion matrix

```{r}
PredictCART = predict(HeartTree, newdata = Test, type = "class")
confusionMatrix = table(Test$DEATH_EVENT, PredictCART)
confusionMatrix
```
Calculate accuracy of CART model:

```{r}
sum(diag(confusionMatrix)) / nrow(Test)
```
## ROC curve

```{r}
library(ROCR)
PredictROC = predict(HeartTree, newdata = Test)
pred = prediction(PredictROC[,2], Test$DEATH_EVENT)
perf = performance(pred, "tpr", "fpr")
plot(perf)
```

## Random Forests

```{r message=FALSE}
library(randomForest)
HeartForest = randomForest(DEATH_EVENT ~ ., data = Train, ntree=200, nodesize=25 )
PredictForest = predict(HeartForest, newdata = Test)
```

Calculate accuracy from confusion matrix:

```{r}
forestMatrix = table(Test$DEATH_EVENT, PredictForest)
sum(diag(forestMatrix)) / nrow(Test)
```
## Cross-validation

```{r message=FALSE}
library(caret)
library(e1071)
numFolds = trainControl( method = "cv", number = 10 )
cpGrid = expand.grid( .cp = seq(0.03,0.06,0.005)) 
train(DEATH_EVENT ~ ., data = Train, method = "rpart", trControl = numFolds, tuneGrid = cpGrid )
```
Calculate accuracy of cross-validated CART-model:

```{r}
HeartTreeCV = rpart(DEATH_EVENT ~ ., data = Train, method="class", cp = 0.045)
PredictCV = predict(HeartTreeCV, newdata = Test, type = "class")
treeMatrixCV = table(Test$DEATH_EVENT, PredictForest)
sum(diag(treeMatrixCV)) / nrow(Test)
```